---
title: "Introduction"
author: "Zebulun Arendsee"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Introduction}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

```{r}
library(onekp)
library(knitr)
library(magrittr)
```

## Accessing the OneKP metadata

All project with the `onekp` R package start at the same place:

```{r}
onekp <- retrieve_onekp()
class(onekp)
```

The `retrieve_onekp` function scrapes the metadata associated with each
transcriptome project from the [1KP public data
page](http://www.onekp.com/public_data.html). It also links each species to its
NCBI taxonomy ID (which is used later to filter by clade).

The only part of the OneKP object that you will need to interact with directly
is the `@table` slot, a data.frame with the form:

```{r, echo = FALSE}
onekp@table %>% head(3) %>% knitr::kable()
```

## Retrieving sequence

To get sequence, first subset the `onekp@table` until it contains only the
species you want. There are several ways to do this.

You can use all the normal tools for subsetting the table directly, e.g.

```{r}
onekp@table <- subset(onekp@table, family == 'Nymphaeaceae')
```

`onekp` also has a few builtin tools for taxonomic selection


```{r}
# filter by species name
filter_by_species(onekp, 'Pinus radiata')

# filter by species NCBI taxon ID
filter_by_species(onekp, 3347)

# filter by clade name scientific name (get all data for the Brassicaceae family)
filter_by_clade(onekp, 'Brassicaceae')

# filter by clade NCBI taxon ID
filter_by_clade(onekp, 3700)
```

Once you have chosen the studies you want, you can retrieve the protein or
transcript FASTA files:

```{r, eval = FALSE}
download_peptides(filter_by_clade(onekp, 'Brassicaceae'))
download_nucleotides(filter_by_clade(onekp, 'Brassicaceae'))
```

This will download the files into a temporary directory. Alternatively, you may
set your own directory with the `dir` argument. The downloaded protein FASTA
files have the extension `.faa` and the DNA files the extension `.fna`. The
basename for each file is the 1KP 4-letter code.
